Quantum internet and quantum machine learning

### 2019-03-05 Nana QML section/QinternetQML.tex
Formatting QML

%\documentclass[aps,prx,twocolumn,preprintnumbers,superscriptaddress]{revtex4}

@@

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{adjustbox}

@@

\begin{document}
%\maketitle

\noindent \textbf{Quantum internet and quantum machine learning}
%\author{Chris, Nana}

@@

\textit{Anomaly detection and fault management---} When there are security breeches in classical data on a network, one desires the ability to predict and detect these disturbances, as well as a method for making one's protocols more robust against these adversaries. Machine learning is often used in anomaly detection and intrusion detection. These algorithms look for unusual data or unusual changes in data. Broadly, there are three classes of anomalies: point, contextual and collective, which refer respectively to single datum anomalies, unusual data with respect to a specified context and clusters of data which together point to an unusual pattern. Both supervised and unsupervised algorithms are used for these settings \cite{thottan2003anomaly, ahmed2007machine}. One of the prime challenges here include determining the presence of an anomaly when little data is available and determining the relevant rate of false positives and false negatives for a particular application.

@@

\textit{Adversarial machine learning---} Machine learning algorithms themselves are vulnerable to security attacks. This comes under the field of adversarial machine learning \cite{huang2011adversarial}. There are two main types of attacks: attacks of the test data (evasion) and attacks of the training data and the machine learning models (poisoning). In real-life deployments of machine learning, data often comes from different sources which makes adversarial attacks more probable. It has been discovered that many machine learning algorithms are in fact vulnerable to adversarial attacks, the first discovered in \cite{szegedy2013intriguing}. A large proportion of the literature then focuses on the details of specific algorithms: the detection of adversaries, their different methods of attack and the particular defenses to those attacks \cite{kurakin2018adversarial}. However, recently, more foundational work has emerged to try to explain the origins of this vulnerability as arising from the high dimensionality of the data involved \cite{goodfellow2014explaining, gilmer2018adversarial,mahloujifar2018curse}.

@@

\textit{Traffic---} Predicting network traffic is becoming increasingly important especially in diverse and complex networks. It is commonly addressed using time-series forecasting (TSF) methods. This can make use of either statistical analysis models or supervised machine learning methods \cite{bermolen2009support, chabaa2010identification, cortez2006internet}]. Non-TSF methods also exist \cite{chen2016predicting, li2016inter}.

@@

\textit{Routing---} Machine learning is most applicable to routing problems where the network is dynamical and thus requires fast updating of new optimal routes. otherwise, already-made routing tables can be sufficient for networks. Since these settings depend on frequent updating on new data, reinforcement learning algorithms are the most appropriate. In particular, Q-learning has performed well in various different networks \cite{wang2006adaptive, forster2007froms, arroyo2007q}.

\textit{Congestion---} Congestion control in a network is also important to ensure the stability of a network and the minimisation of packet loss. Well-known congestion control methods like queue management already exist. However machine learning can be used to enhance the effectiveness of congestion control in various different networks, although mostly for TCP/IP networks \cite{liu2002end, barman2004model, el2005improving}.

@@

\textit{Fully-quantum algorithms---} The first set of these algorithms have chiefly relied on assuming completely quantum devices, keeping coherence throughout the computation and could require full fault-tolerance. For those algorithms claiming up to exponential quantum speedups for supervised algorithms (see \cite{biamonte2017quantum, ciliberto2018quantum}),  the HHL algorithm \cite{harrow2009quantum} for matrix-inversion is often used. However, HHL has a number of drawbacks that make them impractical for near-term quantum devices: (i) the ability to efficiently encode the classical data into quantum states and into quantum memory \cite{aaronson2015read}; (ii) effective read-out of the final quantum state \cite{aaronson2015read}; (iii) generally requiring high circuit-depth and (iv) restrictions on the sparsity and condition numbers of the matrix.

@@

\textit{Hybrid algorithms---} To find algorithms that may be realised in the the near-term, quantum machine learning algorithms are now giving more attention to hybrid classical-quantum algorithms. These algorithms, which include variational methods for optimisation \cite{moll2018quantum}, have short circuit-depths and where the optimisation process is performed iteratively and classically. They are of roughly two types: one that attempts to enhance classical algorithms with classical input data and another where the quantum advantage lies in efficient quantum state preparation, thus using quantum input data. Prominent examples of the former include quantum approximate optimisation algorithm (QAOA) \cite{farhi2014quantum, farhi2016quantum} and the latter include variational quantum eigensolvers (VQE) \cite{peruzzo2014peruzzo, kandala2017hardware}. We return to VQE in the following section as these solve problems for quantum data.

@@

\subsection{Quantum-enhanced machine learning for security and other applications}
\textit{Anomaly detection---} The chief machine learning method for detecting and averting faults and security breaches in classical networks belong to anomaly detection. However, for anomalies in classical data, it appears unlikely that currently available quantum machine learning algorithms can enhance the speed and reliability of detection. One of the primary reasons is the necessity of encoding the classical data into quantum states, which can be very costly \cite{aaronson2015read}. Thus, even if there are quantum-enhanced supervised and unsupervised machine algorithms for anomaly detection in the computational component, the state preparation and read-out demands may be too much. However, the case is different if we begin with quantum data instead and we return to this in section~\ref{sec:sec4}.

\textit{Adversarial quantum machine learning---} Just as machine learning algorithms are vulnerable to attacks, so would quantum-enhanced machine learning algorithms. This is very new area, called adversarial quantum machine learning. Like in adversarial machine learning, the aim is to find more robust quantum machine learning algorithms, and some robust algorithms have been indeed been proposed \cite{wiebe2018hardening}. In addition to finding more robust algorithms, it is also important to understand what the robustness limits of quantum machine learning algorithms actually are, which remains an open problem. A recent result suggests that perhaps just as much quantum resources are necessary for detecting adversaries in higher dimensions as compared to quantum tomography \cite{advql}. Thus it remains unclear the total resource cost of quantum-enhanced machine learning in the presence of adversaries. However, there is a tantalising yet unexplored possibility that perhaps quantum resources can enhance the security of machine learning algorithms, in a similar way that information-theoretic security is afforded to quantum cryptographic protocols.

\textit{Other---} Whether or not there are helpful applications of quantum-enhanced machine learning algorithms to traffic and routing management is currently very unclear and may even appear unlikely. There may be some quantum-enhancements to supervised and unsupervised machine algorithms that could be used in traffic and routing management. However, a key issue still remains in how classical data can be embedded into the relevant quantum states and then read out, in a way that is easier in the quantum setting. The no-cloning theorem forbids reproducing the state and in general overheads in embedding classical information into the relevant quantum states are very high \cite{giovannetti2008quantum, giovannetti2008architectures}. In addition, given the dynamical nature of networks, where machine learning methods appears to be most helpful, the speed of embedding classical data into quantum states must be likewise high. Thus, the necessary quantum resources only to convert classical data to quantum states may overwhelm any computational advantages.

@@

\subsection{Classical machine learning for quantum data}
\textit{Tomography---} Suppose data come naturally in the form of a quantum state or a quantum channel. Then for classical processing of this data over a classical network, the first task is to find its classical description. The canonical methods for this are quantum state tomography and quantum process tomography. However, due to the no-cloning and the cost of quantum measurements, tomography is in general very resource intensive, requiring the number of copies of the quantum states to scale exponentially with the number of qubits per state. However, recent work have revealed methods on efficient state tomography using classical machine learning techniques \cite{Torlai2017, Han2017} over a larger range of states than previously studied.

\textit{Separability---} While tomography reveals the complete classical description of the quantum data, sometimes it may be sufficient to first classify the data in terms of their quantum properties. For instance, methods for classifying quantum states directly in terms of separability have been devised using classical machine learning \cite{Ma2017, Su2017, Gao2018}. Here there are empirical demonstrations of some advantages compared to the CHSH inequality. However, gathering sufficient training states might still remain a problem for states of higher dimensionality.

\textit{Automated experiment design---} In a future quantum internet with entangled quantum networks, it is desirable to find the optimal methods of creating the types and extent of entanglement required with respect to resource constraints. It is also desirable this process may be automated. Recently, such automated methods based on classical reinforcement learning \cite{alexey} have been proposed, to experimentally create a variety of entangled states with more efficiency. This provides an exciting beginning for automated design of future quantum internet protocols.

\textit{VQE---} We saw that variational quantum eigensolvers rely on classical optimisation. When used with input quantum data, they have found success mostly in quantum chemistry \cite{peruzzo2014peruzzo, moll2018quantum}. In the context of quantum networks on the other hand, the most relevant work so far may be its use in quantum data compression \cite{jonromero} which may aid the communication of quantum data over a network. 
%\noindent $\bullet$ Phases of matter for quantum systems processed using machine learning \cite{Juan2017,Torlai2017,Huber2017}.

\subsection{Quantum learning protocols}
\textit{Template matching---} The first quantum algorithms for quantum data most relevant for machine learning are quantum template matching algorithms \cite{saski1, sasaki2}. These are classification algorithms, where each class is represented by a quantum state: a `template'. The task is to find the class to which a given test quantum state belongs, where this state is not identical to any of the template states. It is not clear on its own whether quantum template matching is directly useful in a quantum network setting. However, the ideas introduced here provide the key foundation for supervised learning of quantum data, which can be used in the quantum counterparts to supervised algorithms in traffic prediction, classification and anomaly detection.

\textit{Learning quantum processes---} Suppose one wants to send just enough information about a quantum process over a quantum network in order for the other parties to replicate the use of this process onto any desired quantum state. In the quantum data scenario, one is not a priori given the classical description of this quantum process. Instead, one is provided only a finite number of queries of this quantum process. For a unitary operation, this problem is tackled in \cite{bisio_optimal_2010}, in a problem called the quantum learning of unitary operations. A very interesting observation made here is that the optimal strategy is semi-classical instead of fully-quantum: meaning that it is sufficient for the classical data encoding the estimation of the unknown unitary to be stored. It remains an open question on how these results may change if one extends to more general quantum processes.

\textit{Quantum learning and security---} In a future quantum network with quantum data being exchanged between different parties, it becomes important to detect unusual behaviour in the incoming quantum data. These may be the first signs of a security breach or a fault in the network. For dynamical data in a time-series, for instance, change point detection addresses precisely this problem. This has been extended to the quantum domain \cite{gael1,gael2} where the optimal methods for detecting a change in quantum data are found using methods from state discrimination. For static data, anomaly detection methods based on machine learning become more appropriate as the definition of of unusual behaviour is based on priori training data. Classical anomaly detection algorithms have been applied to quantum data in \cite{sara} for the purpose of error detection, in the case where the classical description for the quantum data is known. However, for cases where the classical description for the quantum data is unknown (as expected over a quantum internet), it is instead much more efficient to directly apply quantum algorithms onto the quantum data directly. Examples of this include several quantum algorithms for anomaly detection was proposed in \cite{liu2018quantum}.

@@

\begin{table}[h]
\centering
\begin{adjustbox}{width=\textwidth}
\begin{tabular}{lllp{5cm}}
Network  & Classical data & Quantum resources  & Quantum data \\
concerns & & and classical data & \\\hhline{|=|=|=|=|}
Individual & Classical   & Quantum-enhanced  & Quantum learning  \\ 
computing & machine learning & machine learning \\ \hline 
Security and faults & Machine learning for anomalies  & Adversarial quantum & Anomaly detection and  \\
& and faults: detection and prediction; & machine learning  & change point detection \\
& Adversarial machine learning &  & for quantum data \\ \hline 
Routing and traffic & Machine learning for \\
& traffic prediction, classification, & Open problems  & Open problems \\
& congestion control and routing\\ \hline 
Distributed  & Distributed machine learning & Distributed quantum & Distributed quantum learning \\
computing & & machine learning\\ \hline 
 Communication & Data compression & Open problems & Data compression and \\
 & and machine learning & & quantum machine learning \\ \hline 
\end{tabular}
\end{adjustbox}
\caption{\label{tab:CommTable2} \textit{Classical and quantum machine learning applications in classical and quantum networks.} Almost all of the categories here are very new and open to exploration in the quantum domain.}
\end{table}

>>>>

@@

%% Language and font encodings
\usepackage[english]{babel}
\usepackage{adjustbox}

@@

\title{Quantum internet and quantum machine learning}

\begin{document}

\maketitle

@@

\subsubsection{Anomaly detection and fault management}

When there are security breeches in classical data on a network, one desires the ability to predict and detect these disturbances, as well as a method for making one's protocols more robust against these adversaries. Machine learning is often used in anomaly detection and intrusion detection. These algorithms look for unusual data or unusual changes in data. Broadly, there are three classes of anomalies: point, contextual and collective, which refer respectively to single datum anomalies, unusual data with respect to a specified context and clusters of data which together point to an unusual pattern. Both supervised and unsupervised algorithms are used for these settings \cite{thottan2003anomaly, ahmed2007machine}. One of the prime challenges here include determining the presence of an anomaly when little data is available and determining the relevant rate of false positives and false negatives for a particular application.

@@

\subsubsection{Adversarial machine learning}

Machine learning algorithms themselves are vulnerable to security attacks. This comes under the field of adversarial machine learning \cite{huang2011adversarial}. There are two main types of attacks: attacks of the test data (evasion) and attacks of the training data and the machine learning models (poisoning). In real-life deployments of machine learning, data often comes from different sources which makes adversarial attacks more probable. It has been discovered that many machine learning algorithms are in fact vulnerable to adversarial attacks, the first discovered in \cite{szegedy2013intriguing}. A large proportion of the literature then focuses on the details of specific algorithms: the detection of adversaries, their different methods of attack and the particular defenses to those attacks \cite{kurakin2018adversarial}. However, recently, more foundational work has emerged to try to explain the origins of this vulnerability as arising from the high dimensionality of the data involved \cite{goodfellow2014explaining, gilmer2018adversarial,mahloujifar2018curse}.

@@

\subsubsection{Traffic}

Predicting network traffic is becoming increasingly important especially in diverse and complex networks. It is commonly addressed using time-series forecasting (TSF) methods. This can make use of either statistical analysis models or supervised machine learning methods \cite{bermolen2009support, chabaa2010identification, cortez2006internet}]. Non-TSF methods also exist \cite{chen2016predicting, li2016inter}.

@@

\subsubsection{Routing}

Machine learning is most applicable to routing problems where the network is dynamical and thus requires fast updating of new optimal routes. otherwise, already-made routing tables can be sufficient for networks. Since these settings depend on frequent updating on new data, reinforcement learning algorithms are the most appropriate. In particular, Q-learning has performed well in various different networks \cite{wang2006adaptive, forster2007froms, arroyo2007q}.

\subsubsection{Congestion}

Congestion control in a network is also important to ensure the stability of a network and the minimisation of packet loss. Well-known congestion control methods like queue management already exist. However machine learning can be used to enhance the effectiveness of congestion control in various different networks, although mostly for TCP/IP networks \cite{liu2002end, barman2004model, el2005improving}.

@@

\subsubsection{Fully-quantum algorithms}

The first set of these algorithms have chiefly relied on assuming completely quantum devices, keeping coherence throughout the computation and could require full fault-tolerance. For those algorithms claiming up to exponential quantum speedups for supervised algorithms (see \cite{biamonte2017quantum, ciliberto2018quantum}),  the HHL algorithm \cite{harrow2009quantum} for matrix-inversion is often used. However, HHL has a number of drawbacks that make them impractical for near-term quantum devices: (i) the ability to efficiently encode the classical data into quantum states and into quantum memory \cite{aaronson2015read}; (ii) effective read-out of the final quantum state \cite{aaronson2015read}; (iii) generally requiring high circuit-depth and (iv) restrictions on the sparsity and condition numbers of the matrix.

@@

\subsubsection{Hybrid algorithms}

To find algorithms that may be realised in the the near-term, quantum machine learning algorithms are now giving more attention to hybrid classical-quantum algorithms. These algorithms, which include variational methods for optimisation \cite{moll2018quantum}, have short circuit-depths and where the optimisation process is performed iteratively and classically. They are of roughly two types: one that attempts to enhance classical algorithms with classical input data and another where the quantum advantage lies in efficient quantum state preparation, thus using quantum input data. Prominent examples of the former include quantum approximate optimisation algorithm (QAOA) \cite{farhi2014quantum, farhi2016quantum} and the latter include variational quantum eigensolvers (VQE) \cite{peruzzo2014peruzzo, kandala2017hardware}. We return to VQE in the following section as these solve problems for quantum data.

@@

\subsection{Quantum-enhanced machine learning for security and other applications}

\subsubsection{Anomaly detection}

The chief machine learning method for detecting and averting faults and security breaches in classical networks belong to anomaly detection. However, for anomalies in classical data, it appears unlikely that currently available quantum machine learning algorithms can enhance the speed and reliability of detection. One of the primary reasons is the necessity of encoding the classical data into quantum states, which can be very costly \cite{aaronson2015read}. Thus, even if there are quantum-enhanced supervised and unsupervised machine algorithms for anomaly detection in the computational component, the state preparation and read-out demands may be too much. However, the case is different if we begin with quantum data instead and we return to this in section~\ref{sec:sec4}.

\subsubsection{Adversarial quantum machine learning}

Just as machine learning algorithms are vulnerable to attacks, so would quantum-enhanced machine learning algorithms. This is very new area, called adversarial quantum machine learning. Like in adversarial machine learning, the aim is to find more robust quantum machine learning algorithms, and some robust algorithms have been indeed been proposed \cite{wiebe2018hardening}. In addition to finding more robust algorithms, it is also important to understand what the robustness limits of quantum machine learning algorithms actually are, which remains an open problem. A recent result suggests that perhaps just as much quantum resources are necessary for detecting adversaries in higher dimensions as compared to quantum tomography \cite{advql}. Thus it remains unclear the total resource cost of quantum-enhanced machine learning in the presence of adversaries. However, there is a tantalising yet unexplored possibility that perhaps quantum resources can enhance the security of machine learning algorithms, in a similar way that information-theoretic security is afforded to quantum cryptographic protocols.

\subsubsection{Other}

Whether or not there are helpful applications of quantum-enhanced machine learning algorithms to traffic and routing management is currently very unclear and may even appear unlikely. There may be some quantum-enhancements to supervised and unsupervised machine algorithms that could be used in traffic and routing management. However, a key issue still remains in how classical data can be embedded into the relevant quantum states and then read out, in a way that is easier in the quantum setting. The no-cloning theorem forbids reproducing the state and in general overheads in embedding classical information into the relevant quantum states are very high \cite{giovannetti2008quantum, giovannetti2008architectures}. In addition, given the dynamical nature of networks, where machine learning methods appears to be most helpful, the speed of embedding classical data into quantum states must be likewise high. Thus, the necessary quantum resources only to convert classical data to quantum states may overwhelm any computational advantages.

@@

\subsection{Classical machine learning for quantum data}

\subsubsection{Tomography}

Suppose data come naturally in the form of a quantum state or a quantum channel. Then for classical processing of this data over a classical network, the first task is to find its classical description. The canonical methods for this are quantum state tomography and quantum process tomography. However, due to the no-cloning and the cost of quantum measurements, tomography is in general very resource intensive, requiring the number of copies of the quantum states to scale exponentially with the number of qubits per state. However, recent work have revealed methods on efficient state tomography using classical machine learning techniques \cite{Torlai2017, Han2017} over a larger range of states than previously studied.

\subsubsection{Separability}

While tomography reveals the complete classical description of the quantum data, sometimes it may be sufficient to first classify the data in terms of their quantum properties. For instance, methods for classifying quantum states directly in terms of separability have been devised using classical machine learning \cite{Ma2017, Su2017, Gao2018}. Here there are empirical demonstrations of some advantages compared to the CHSH inequality. However, gathering sufficient training states might still remain a problem for states of higher dimensionality.

\subsubsection{Automated experiment design}

In a future quantum internet with entangled quantum networks, it is desirable to find the optimal methods of creating the types and extent of entanglement required with respect to resource constraints. It is also desirable this process may be automated. Recently, such automated methods based on classical reinforcement learning \cite{alexey} have been proposed, to experimentally create a variety of entangled states with more efficiency. This provides an exciting beginning for automated design of future quantum internet protocols.

\subsubsection{VQE}

We saw that variational quantum eigensolvers rely on classical optimisation. When used with input quantum data, they have found success mostly in quantum chemistry \cite{peruzzo2014peruzzo, moll2018quantum}. In the context of quantum networks on the other hand, the most relevant work so far may be its use in quantum data compression \cite{jonromero} which may aid the communication of quantum data over a network. 
%\noindent $\bullet$ Phases of matter for quantum systems processed using machine learning \cite{Juan2017,Torlai2017,Huber2017}.

\subsection{Quantum learning protocols}

\subsubsection{Template matching}

The first quantum algorithms for quantum data most relevant for machine learning are quantum template matching algorithms \cite{saski1, sasaki2}. These are classification algorithms, where each class is represented by a quantum state: a `template'. The task is to find the class to which a given test quantum state belongs, where this state is not identical to any of the template states. It is not clear on its own whether quantum template matching is directly useful in a quantum network setting. However, the ideas introduced here provide the key foundation for supervised learning of quantum data, which can be used in the quantum counterparts to supervised algorithms in traffic prediction, classification and anomaly detection.

\subsubsection{Learning quantum processes}

Suppose one wants to send just enough information about a quantum process over a quantum network in order for the other parties to replicate the use of this process onto any desired quantum state. In the quantum data scenario, one is not a priori given the classical description of this quantum process. Instead, one is provided only a finite number of queries of this quantum process. For a unitary operation, this problem is tackled in \cite{bisio_optimal_2010}, in a problem called the quantum learning of unitary operations. A very interesting observation made here is that the optimal strategy is semi-classical instead of fully-quantum: meaning that it is sufficient for the classical data encoding the estimation of the unknown unitary to be stored. It remains an open question on how these results may change if one extends to more general quantum processes.

\subsubsection{Quantum learning and security}

In a future quantum network with quantum data being exchanged between different parties, it becomes important to detect unusual behaviour in the incoming quantum data. These may be the first signs of a security breach or a fault in the network. For dynamical data in a time-series, for instance, change point detection addresses precisely this problem. This has been extended to the quantum domain \cite{gael1,gael2} where the optimal methods for detecting a change in quantum data are found using methods from state discrimination. For static data, anomaly detection methods based on machine learning become more appropriate as the definition of of unusual behaviour is based on priori training data. Classical anomaly detection algorithms have been applied to quantum data in \cite{sara} for the purpose of error detection, in the case where the classical description for the quantum data is known. However, for cases where the classical description for the quantum data is unknown (as expected over a quantum internet), it is instead much more efficient to directly apply quantum algorithms onto the quantum data directly. Examples of this include several quantum algorithms for anomaly detection was proposed in \cite{liu2018quantum}.

@@

\begin{table}[h]
\centering
\begin{adjustbox}{width=\textwidth}
\begin{tabular}{lllp{5cm}}
Network  & Classical data & Quantum resources  & Quantum data \\
concerns & & and classical data & \\\hhline{|=|=|=|=|}
Individual & Classical   & Quantum-enhanced  & Quantum learning  \\ 
computing & machine learning & machine learning \\ \hline 
Security and faults & Machine learning for anomalies  & Adversarial quantum & Anomaly detection and  \\
& and faults: detection and prediction; & machine learning  & change point detection \\
& Adversarial machine learning &  & for quantum data \\ \hline 
Routing and traffic & Machine learning for \\
& traffic prediction, classification, & Open problems  & Open problems \\
& congestion control and routing\\ \hline 
Distributed  & Distributed machine learning & Distributed quantum & Distributed quantum learning \\
computing & & machine learning\\ \hline 
 Communication & Data compression & Open problems & Data compression and \\
 & and machine learning & & quantum machine learning \\ \hline 
\end{tabular}
\end{adjustbox}
\caption{\label{tab:CommTable2} Classical and quantum machine learning applications in classical and quantum networks. Almost all of the categories here are very new and open to exploration in the quantum domain.}
\end{table}
